Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-04-04 manuscript/manuscript.tex
Modifications on introduction in manuscript

\usepackage[utf8]{inputenc}
\usepackage{timet}
\usepackage{amsmath}
\usepackage{graphicx}
%~ \usepackage{todonotes} % to make annotations on margins
\usepackage[disable]{todonotes} % to disable annotations on margins

@@

In order to develop a forward model that computes the gravity field generated 
by any tesseroid with arbitrary variable density, the Taylor series expansion 
is not well suited, because it would need the series expansion terms for each 
density function.
On the other hand, the GLQ consists in approximating the integral by a 
weighted sum of the effect of point masses located at scaled nodes of the 
Legendre polynomials, where the density variation is summarised into the 
weights.
Thus, this kind of numerical approximation allows to calculate the gravity 
fields for any density function without any other information but the function 
itself.

@@

\citet{Uieda2016} developed a forward gravity model based on tesseroids with 
homogeneous densities using the GLQ approximation method.
It has been originally implemented through command line programs written in C 
programming language.
However, it has been ultimately ported to Python and included in the 
open-source library Fatiando a Terra \citep{Uieda2013} for geophysical 
modelling and inversion.
\todo{hace falta aclarar sobre el software de leo?}

\citet{Ku1977} proved that the GLQ becomes less accurate when the computation 
point is closer to the tesseroid or for smaller GLQ order.
Although the model developed by \citet{Uieda2016} uses a second order GLQ, it 
ensures the accuracy implementing a modified version of the adaptive 
discretization of \citet{Li2011}.
It consist in splitting the tesseroid into smaller ones when a certain 
distance-size ratio is lower than a predefined value ($D$) that controls the 
accuracy of the computation.
\citet{Uieda2016} have also objectively obtained standard values of $D$ for 
gravity potential, gradients and tensor components comparing the numerical 
model with the fields generated by a spherical shell, which constitutes a 
special case of the volume integrals that has analytical solution 
\citep{LaFehr1991, Mikuska2006, Grombein2013}
\todo{citar mas??}.

In addition, a varying density inside an arbitrary tesseroid introduces a new 
kind of error into the numerical approximation. This encouraged us to develop 
a new density-based discretization algorithm that divides the tesseroid into 
smaller ones taking into account the variation of the density function.
The number of subdivisions and thus the accuracy of the approximation will be 
controlled by a delta ratio $\delta$, predefined by the user.

We have enhanced the implementation of \citet{Uieda2016} with a new forward 
gravity model that allows to compute the gravity fields generated by any 
tesseroid with an arbitrary density function on any external point.
It is based on the GLQ approximation, makes use of the modified distance-size 
discretization and the new density-based discretization algorithms that 
guarantee the accuracy even for high varying density functions.
It is written in Python and Cython languages, constituting an easy to use 
library that runs precompiled code for the more time consuming functions.
It also makes use of some previously existing classes and functions from 
Fatiando a Terra, what saved us time and allowed us to focus on the new 
enhancements.
Furthermore, we have the intention to include it into a future release of the 
Fatiando a Terra library in order to make its distribution and maintenance 
easier.

In order to ensure the accuracy of the numerical approximation, we have 
conducted similar tests as the one performed by \citet{Uieda2016} by comparing 
the results of the numerical model with the analytical solutions for a 
spherical shell with linear, exponential and discontinuous density.
Furthermore we had to derive the analytical expressions for these density 
functions.

>>>>

\usepackage[utf8]{inputenc}
\usepackage{timet}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{todonotes} % to make annotations on margins
%~ \usepackage[disable]{todonotes} % to disable annotations on margins

@@

In order to develop a forward model that computes the gravitational 
field generated by any tesseroid with arbitrary variable density, the 
Taylor series expansion is not well suited, because it would need the 
series expansion terms for each density function.
On the other hand, the GLQ consists in approximating the integral by a 
weighted sum of the effect of point masses located at scaled nodes of the 
Legendre polynomials, where the density variation is summarised into the 
weights.
Thus, this kind of numerical approximation allows to calculate the 
gravitational fields for any density function without any other 
information but the function itself.

@@

\citet{Uieda2016} developed a forward gravitational model based on 
tesseroids with homogeneous densities using the GLQ approximation 
method.
The main issue of the GLQ is that the approximation becomes less 
accurate when the computation point is closer to the tesseroid of for 
smaler GLQ order \citep{Ku1977}.
Although the model developed by \citet{Uieda2016} uses a second order GLQ, it 
ensures the accuracy implementing a modified version of the adaptive 
discretization of \citet{Li2011}.
It consist in splitting the tesseroid into smaller ones when a certain 
distance-size ratio is lower than a predefined value ($D$) that controls the 
accuracy of the computation.
\citet{Uieda2016} have also objectively obtained standard values of $D$ 
for gravitational potential, gradients and tensor components comparing 
the numerical model with the fields generated by a spherical shell, 
which constitutes a special case of the volume integrals that has 
analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013} 
\todo{citar mas??}.

On the other hand, a varying density inside an arbitrary tesseroid 
introduces a new kind of error into the numerical approximation.
This encouraged us to develop a new density-based discretization 
algorithm that divides the tesseroid into smaller ones taking into 
account the variation of the density function.
The number of subdivisions and thus the accuracy of the approximation 
will be controlled by predefined delta ratio $\delta$.

In summary, we present a new forward gravitational model that allows to 
compute the gravitational fields generated by any tesseroid with an 
arbitrary continuous density function on any external point.
It is based on the GLQ approximation whose accuracy is controlled by 
the modified distance-size discretization \citep{Uieda2016} 
and the new density-based discretization algorithms.

In order to ensure the accuracy of the numerical approximation we have 
determined predefined values for the $D$ and $\delta$ parameters by 
comparing the numerical approximation with the analytical solutions for 
a spherical shell with linear and exponential density, which had to be 
derived.

Furthermore, we developed a Python library that implements this forward 
model.
I makes use of some previously existing classes and functions 
from the Python library Fatiando a Terra for geophysical modelling, 
what saved us time and allowed us to focus on the new model.

Furthermore, we have the intention to include it into a future release 
of the Fatiando a Terra library in order to make its distribution and 
maintenance easier.
\todo{include this?}
